Stay in shape! Homework exam II

### 2016-04-04 geometricAlgorithms/homeworkExam/2/homeworkExam.tex
imporved the proofs of 1, also made psuedocode fit.

			\If{$h_i$ is parallel to the x-axis}
				\State
				\Return $False$
			\EndIf
			\State
			\Comment{Check where the x axes crossings are, and write it into
			a variable pair, if both pairs are filled its not case ii.}
			\If{$h_i$ points left}
			\State
				$b_r \gets h_i$
			\EndIf
			\If{$h_i$ points right}
			\State
				$b_l \gets h_i$
			\EndIf
			\If{$b_r$ not equal to $None$ and $b_l$ not equal to $None$}
				\State
				\Return $False$
			\EndIf
		\EndFor
		\State
		\Return $True$
	\EndFunction
\end{algorithmic}
\subsection{Formulate and proof the RIC lemma}
\paragraph{Lemma}
Let $1 \leq i \leq n$, let $C_i$ be the defined as above and let $d_i$ be
the largest disk in $C_i$. Then we have
\begin{enumerate}[(i)]
	\item If $d_{i-1} \subset h_i,$ $then$ $d_i = d_{i-1}$
	\item If $d_{i-1} \not\subset h_{i},$ $then$ $C_i$ does not intersect the x-axis
		or $b_i \cap l_i \neq \emptyset$ where $b_i$ is the boundary of $d_i$ and $l_i$
		is the line bounding $h_i$.
\end{enumerate}
\paragraph{Proof}
\begin{enumerate}[(i)]
	\item Let $d_{i-1} \subset h_i$. Because $C_i = C_{i-1}\cap h_i$ and
		$d_{i-1} \subset C_{i-1}$ this means that $d_{i-1} \subset C_i$.
		Furthermore,the optimal disk in $C_i$
		cannot be better than the optimal disk in $C_{i-1}$, since
		$C_i \subseteq C_{i-1}$. Hence, $d_{i-1}$ is the optimal disk
		in $C_i$ as well.
%TODO: replace with using a line between center(d_{i-1}) and
%		center(d_i) then reason about how d_i will keep growing
%		 until it hits h_i.
	\item Let $d_{i-1} \not\subset h_{i}$. Suppose for a contradiction that
		$C_i$ intersects the x-axis and that $b_i \cap l_i = \emptyset$.
		Consider the line segment $s$ between $center(d_i)$ and 
		$center(d_{i-1})$.
		Where $center$ is a function that returns the center vertex of a disk.
		We have $d_{i-1} \subset C_{i-1}$ and
		since $C_i \subseteq C_{i-1}$, also $d_i \subset C_{i-1}$.
		Together with the convexity of $C_{i-1}$, this implies that
		the line segment $s$ is contained within $C_{i-1}$.
		Since $d_{i-1}$ is the optimal disk in $C_{i-1}$ and the optimum is to
		maximize $r_i$ the radius of $d_i$. Consider $d_i$
		where $s$ is maximized, now to increase $r_i$ we move $d_i$ along $s$,
		we can keep doing this until $b_i \cap h_i$.
		So $d_i$ needs to intersect $h_i$ otherwise $d_i$ is not the optimum
		disk in $C_i$. This shows it cannot be the case that $C_i$ intersects
		the x-axis and $b_i\cap l_i = \emptyset$.

@@

We will prove that a$calculateBiggestDisk(H)$ will run in $\Theta(n)$ expected
time with help of backwards analyses: Consider the situation after $h_i$ is
inserted, and $d_i$ is computed (either by $d_i = d_{i-1}$ or with help
of $calculateBiggestDiskFrom(h_i,H)$). If $h_i$ does not bound the feasible
region, or $d_i \subset h_i$, then the addition step was cheap
and took $\Theta(1)$ time. But if $d_i \not\subset h_i$ then the addition
step was expensive.

@@

@@

		A more crude way to put it is that you can have multiple optimum
		that don't intersect the same points.
	\item \emph{The optimum must already be defined by O(1) of the input
		objects.}
		I think this is true, you can use some math to figure out a new
		optimum.
	\item \emph{The analysis must work out.}
		It doesn't work out because you can't guarantee that you've just
		found a local optimum instead of a global optimum. I mean you
		could use this as a heuristic but not as an answer.
\end{itemize}
\section{A data structure for horizontal line segments}
\subsection{Analyze the storage requirements}
As can be read from lemma 10.2 (and the suceeding proof) from
the book \cite{book}, an interval tree uses $O(n)$ storage.

@@

\subsection{Give pseudo code for the query algorithm}
\begin{algorithmic}[1]
	\Function{select}{$v$, $x \times [y:y']$}
		\If{$v$ is a leaf and $y < y_{val}(v) < y'$}
			\State
			\Return \Call{selectRange}{$subtree(v), x$}
		\EndIf
		\If{$y' < y_{val}(v)$}
		\State
		\Return \Call{select}{$c_{left}(v), x \times [y:y']$}
		\EndIf
		\If{$y > y_{val}(v)$}
		\State
		\Return \Call{select}{$c_{right}(v), x \times [y:y']$}
		\EndIf
		\State
		\Comment{If we get here it means we have a split node}
		\State
		\Return $\Call{reportLeft}{v, x\times y, c_{right}}\cup
		\Call{reportRight}{v,x\times y', c_{left}}$
	\EndFunction
	\algstore{tree}
\end{algorithmic}
\newpage
\begin{algorithmic}[1]
	\algrestore{tree}
	\Function{reportRight}{$v$, $x \times y$}
		\State
		$\phi \gets y \leq y_{val}(v)$
		\If{$isLeaf(v)$}
			\If{$\phi$}
				\State
				\Return $\Call{selectRange}{subtree(v),x}$
			\EndIf
			\State
			\Return $\emptyset$
		\EndIf
		\If{$\phi$}
		\State
		\Return $\Call{reportRight}{c_{left}(v), x \times y} \cup
		\Call{selectRange}{subtree(c_{right}(v)),x}$
		\EndIf
		\State
		\Return $\Call{reportRight}{c_{right}(v), x \times y}$
	\EndFunction
	\Function{reportLeft}{$v$, $x \times y$}
		\State
		$\phi \gets y > y_{val}(v)$
		\If{$isLeaf(v)$}
			\If{$\phi$}
				\State
				\Return $\Call{selectRange}{subtree(v),x}$
			\EndIf
			\State
			\Return $\emptyset$
		\EndIf
		\If{$\phi$}
		\State
		\Return $\Call{reportLeft}{c_{right}(v), x \times y} \cup
		\Call{selectRange}{subtree(c_{left}(v)),x}$
		\EndIf
		\State
		\Return $\Call{reportLeft}{c_{left}(v), x \times y}$
	\EndFunction
	\Function{selectRange}{$v$, $x$}
	\If{$v$ is a leaf}
		\State
		\Return $\emptyset$
	\EndIf
	\If{$x < x_{mid}(v)$}
		\State
		$r \gets $ starting at the leftmost endpoint keep
		returning all the intervals in $l_{left}(v)$ until the
		intervals no longer contain $x$. (basically a takeWhile
		operation in your favorite language).
		\State
		\Return $r \cup \Call{selectRange}{c_{left}(v),x}$
	\EndIf
	\State
	$r \gets $ starting at the rightmost endpoint keep
	returning all the intervals in $l_{right}(v)$ until the
	intervals no longer contain $x$.
	\State
	\Return $r \cup \Call{selectRange}{c_{right}(v),x}$
	\EndFunction
\end{algorithmic}
\subsection{Analyze the query time}

>>>>

			\If{$h_i$ is parallel to the x-axis}
				\State
				\Return $False$
			\EndIf
			\State
			\Comment{Check where the x axes crossings are, and write it into
			a variable pair, if both pairs are filled its not case ii.}
			\If{$h_i$ points left}
			\State
				$b_r \gets h_i$
			\EndIf
			\If{$h_i$ points right}
			\State
				$b_l \gets h_i$
			\EndIf
			\If{$b_r$ not equal to $None$ and $b_l$ not equal to $None$}
				\State
				\Return $False$
			\EndIf
		\EndFor
		\State
		\Return $True$
	\EndFunction
\end{algorithmic}
\subsection{Formulate and proof the RIC lemma}
\paragraph{Lemma}
Let $1 \leq i \leq n$, let $C_i$ be the defined as above and let $d_i$ be
the largest disk in $C_i$. Then we have
\begin{enumerate}[(i)]
	\item If $d_{i-1} \subset h_i,$ $then$ $d_i = d_{i-1}$
	\item If $d_{i-1} \not\subset h_{i},$ $then$ $C_i$ does not intersect the x-axis
		or $b_i \cap l_i \neq \emptyset$ where $b_i$ is the boundary of $d_i$ and $l_i$
		is the line bounding $h_i$.
\end{enumerate}
\paragraph{Proof}
\begin{enumerate}[(i)]
	\item Let $d_{i-1} \subset h_i$. Because $C_i = C_{i-1}\cap h_i$ and
		$d_{i-1} \subset C_{i-1}$ this means that $d_{i-1} \subset C_i$.
		Furthermore,the optimal disk in $C_i$
		cannot be better than the optimal disk in $C_{i-1}$, since
		$C_i \subseteq C_{i-1}$. Hence, $d_{i-1}$ is the optimal disk
		in $C_i$ as well.
	\item Let $d_{i-1} \not\subset h_{i}$. Suppose for a contradiction that
		$C_i$ intersects the x-axis and that $b_i \cap l_i = \emptyset$.
		Consider the line segment $s$ between $center(d_i)$ and 
		$center(d_{i-1})$.
		Where $center$ is a function that returns the center vertex of a disk.
		We have $d_{i-1} \subset C_{i-1}$ and
		since $C_i \subseteq C_{i-1}$, also $d_i \subset C_{i-1}$.
		Together with the convexity of $C_{i-1}$, this implies that
		the line segment $s$ is contained within $C_{i-1}$.
		Consider $d_i$ with the center at the end of $s$.
		This is the smallest possible disk,
		now to increase the radius we move $d_i$ along $s$,
		we can keep doing this until $b_i \cap h_i$.
		$d_i$ needs to intersect $h_i$ otherwise $d_i$ is not the optimum
		disk in $C_i$. This shows it cannot be the case that $C_i$ intersects
		the x-axis and $b_i\cap l_i = \emptyset$.

@@

We will prove that $calculateBiggestDisk(H)$ will run in $\Theta(n)$ expected
time with help of backwards analyses: Consider the situation after $h_i$ is
inserted, and $d_i$ is computed (either by $d_i = d_{i-1}$ or with help
of $calculateBiggestDiskFrom(h_i,H)$). If $h_i$ does not bound the feasible
region, or $d_i \subset h_i$, then the addition step was cheap
and took $\Theta(1)$ time. But if $d_i \not\subset h_i$ then the addition
step was expensive.

@@

One $\Theta(n)$ for dealing with case ii, one $\Theta(n)$ for shuffling and
the final $\Theta(n)$ for the algorithm.

@@

		A more crude way to put it is that you can have multiple optimum
		that don't intersect the same points.
	\item \emph{The optimum must already be defined by O(1) of the input
		objects.}
		I think this is true, you can use some math to figure out a new
		optimum.
	\item \emph{The analysis must work out.}
		It doesn't work out because you can't guarantee that you've just
		found a local optimum instead of a global optimum. I mean you
		could use this as a heuristic but not as an answer.
\end{itemize}
\section{A data structure for horizontal line segments}
\subsection{Analyze the storage requirements}
As can be read from lemma 10.2 (and the succeeding proof) from
the book \cite{book}, an interval tree uses $O(n)$ storage.

@@

\newpage
\subsection{Give pseudo code for the query algorithm}
\begin{algorithmic}[1]
	\Function{select}{$v$, $x \times [y:y']$}
		\If{$v$ is a leaf and $y < y_{val}(v) < y'$}
			\State
			\Return \Call{selectRange}{$subtree(v), x$}
		\EndIf
		\If{$y' < y_{val}(v)$}
		\State
		\Return \Call{select}{$c_{left}(v), x \times [y:y']$}
		\EndIf
		\If{$y > y_{val}(v)$}
		\State
		\Return \Call{select}{$c_{right}(v), x \times [y:y']$}
		\EndIf
		\State
		\Comment{If we get here it means we have a split node}
		\State
		\Return $\Call{reportLeft}{v, x\times y, c_{right}}\cup
		\Call{reportRight}{v,x\times y', c_{left}}$
	\EndFunction
	\Function{reportRight}{$v$, $x \times y$}
		\State
		$\phi \gets y \leq y_{val}(v)$
		\If{$isLeaf(v)$}
			\If{$\phi$}
				\State
				\Return $\Call{selectRange}{subtree(v),x}$
			\EndIf
			\State
			\Return $\emptyset$
		\EndIf
		\If{$\phi$}
		\State
		\Return $\Call{reportRight}{c_{left}(v), x \times y} \cup
		\Call{selectRange}{subtree(c_{right}(v)),x}$
		\EndIf
		\State
		\Return $\Call{reportRight}{c_{right}(v), x \times y}$
	\EndFunction
	\Function{reportLeft}{$v$, $x \times y$}
		\State
		$\phi \gets y > y_{val}(v)$
		\If{$isLeaf(v)$}
			\If{$\phi$}
				\State
				\Return $\Call{selectRange}{subtree(v),x}$
			\EndIf
			\State
			\Return $\emptyset$
		\EndIf
		\If{$\phi$}
		\State
		\Return $\Call{reportLeft}{c_{right}(v), x \times y} \cup
		\Call{selectRange}{subtree(c_{left}(v)),x}$
		\EndIf
		\State
		\Return $\Call{reportLeft}{c_{left}(v), x \times y}$
	\EndFunction
	\algstore{tree}
\end{algorithmic}
\newpage
\begin{algorithmic}[1]
	\algrestore{tree}
	\Function{selectRange}{$v$, $x$}
	\If{$v$ is a leaf}
		\State
		\Return $\emptyset$
	\EndIf
	\If{$x < x_{mid}(v)$}
		\State
		$r \gets $ starting at the leftmost endpoint keep
		returning all the intervals in $l_{left}(v)$ until the
		intervals no longer contain $x$. (basically a takeWhile
		operation in your favorite language).
		\State
		\Return $r \cup \Call{selectRange}{c_{left}(v),x}$
	\EndIf
	\State
	$r \gets $ starting at the rightmost endpoint keep
	returning all the intervals in $l_{right}(v)$ until the
	intervals no longer contain $x$.
	\State
	\Return $r \cup \Call{selectRange}{c_{right}(v),x}$
	\EndFunction
\end{algorithmic}
\subsection{Analyze the query time}
